---
title: "tydy data and other datasets"
author: "Yi"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# load packages

load relevant packages and limit the number of lines printed in a tibble.

```{r}
library(tidyverse)
options(tibble.print_min = 5)
```

# load data

```{r}
pulse_df = haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names()
pulse_df
```

# `pivot_longer`

transpose from wide to long

```{r}
pulse_tidy_data = 
  pivot_longer(
    pulse_df,
    bdi_score_bl:bdi_score_12m,
    names_to = "visit",
    values_to = "bdi"
  )

pulse_tidy_data
```
remove prefix

```{r}
pulse_tidy_data = 
  pivot_longer(
    pulse_df, 
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi")

pulse_tidy_data
```

combined
changing bl to 00m for consistency across visits and converting visit to a factor variable. (It’s possible that you would want visit to be a numeric variable instead, which could be done with a different call to mutate.) Lastly, it’s nice to organize the data into a reasonable order.

```{r}
pulse_df = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") %>%
  janitor::clean_names() %>%
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") %>%
  relocate(visit) %>%
  mutate(
    visit = replace(visit, visit == "bl", "00m"),
    visit = factor(visit)) %>%
  arrange(id, visit)

print(pulse_df, n = 12)
```

# Exercise

```{r}
litters_df = 
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  select(litter_number, ends_with("weight")) %>% 
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = "gd",
    values_to = "weight") %>%
  mutate(gd = recode(gd, "gd0_weight" = 0, "gd18_weight" = 18))

print(litters_df)
```
# `pivot_wider`

```{r}
analysis_result = tibble(
  group = c("treatment", "treatment", "placebo", "placebo"),
  time = c("pre", "post", "pre", "post"),
  mean = c(4, 8, 3.5, 4)
)

analysis_result
```

transpose from long to wide
use knitr::kable() to produce a nicer table for reading

```{r}
pivot_wider(
  analysis_result, 
  names_from = "time", 
  values_from = "mean") %>% 
  knitr::kable()
```


# Binding rows


read the data and add a variable to each dataframe indicating the movie

```{r}
fellowship_ring = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") %>%
  mutate(movie = "fellowship_ring")

two_towers = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "F3:H6") %>%
  mutate(movie = "two_towers")

return_king = 
  readxl::read_excel("./data/LotR_Words.xlsx", range = "J3:L6") %>%
  mutate(movie = "return_king")
```

stack up the data frames
make it longer
mutate the race to lower case
reorder the columns 

```{r}
lotr_tidy = 
  bind_rows(fellowship_ring, two_towers, return_king) %>%
  janitor::clean_names() %>%
  pivot_longer(
    female:male,
    names_to = "gender", 
    values_to = "words") %>%
  mutate(race = str_to_lower(race)) %>% 
  select(movie, everything()) 

lotr_tidy
```

# Joining datasets

There are four major ways join data frames x and y:

*   Inner: keeps data that appear in both x and y
*   Left: keeps data that appear in x
*   Right: keeps data that appear in y
*   Full: keeps data that appear in either x or y


use \`1\` to tell R to look for a number

```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv") %>%
  janitor::clean_names() %>%
  mutate(
    sex = recode(sex, `1` = "male", `2` = "female"),
    sex = factor(sex))

litter_data = 
  read_csv("./data/FAS_litters.csv") %>%
  janitor::clean_names() %>%
  separate(group, into = c("dose", "day_of_tx"), sep = 3) %>%
  relocate(litter_number) %>%
  mutate(
    wt_gain = gd18_weight - gd0_weight,
    dose = str_to_lower(dose))

fas_data = 
  left_join(pup_data, litter_data, by = "litter_number")

fas_data
```
